---
title: "Week 7, Day 1"
output: html_document
---

```{r setup, include=FALSE}
# We need the PPBDS.data package because it includes the qscores data which we
# will use for this exercise. rsample is a package from the tidymodels
# collection of packages which makes random sample creation easier. See The
# Primer for examples on its use.

knitr::opts_chunk$set(echo = FALSE)
library(PPBDS.data)
library(rsample)
library(tidyverse)
```

We learned how to estimate a posterior probability distribution for a single parameter in Chapter 6. But, in doing so, we had to do a lot of work, calculating the joint distribution 
$p(models, data)$ by hand, and then deriving the conditional probability. What a bother! Fortunately, the bootstrap provides an easier and more generalizable approach.

Define the parameter H as the average number of hours of work reported by students for college courses. 


## Scene 1

**Prompt:** Create a tibble called `q_resamples` with three columns: `splits`, `id` and `boot`.  The first two columns are created automatically when you use `rsample::bootstraps()`. The last variable is a list column which includes a tibble with resampled data from `qscores`. There only column you need to keep from `qscores` is `hours`. Refer to Chapter 7 in the *The Primer* for hints.



```{r sc1}
set.seed(9)
q_resamples <- qscores %>%
  select(hours) %>% 
  bootstraps(times = 1000) %>%
  mutate(boot = map(splits, ~ analysis(.)))
```


## Scene 2

**Prompt:** Add two more columns to `q_resamples`: `hours_sampled` (the vector of hours pulled from each row of `boot`) and `mean_hours` (the mean of each `hours_sampled`). 


```{r sc2}
ql_resamples <- qscores %>%
  select(hours) %>% 
  bootstraps(times = 1000) %>%
  mutate(boot = map(splits, ~ analysis(.))) %>%
  mutate(hours_sampled = map(boot, ~ pull(., hours))) %>% 
  mutate(mean_hours = map_dbl(hours_sampled, ~ mean(.)))

```

## Scene 3

**Prompt:** Create a graphic of the posterior probability distribution for H. Interpret it.

```{r}
ql_resamples %>% 
  ggplot(aes(x = mean_hours, 
             y = after_stat(count/sum(count)))) +
    geom_histogram(binwidth = 0.02) +
    labs(x = "Hours", 
         y = "Probability",
         title = "Posterior Distribution for the Average Hours")
```


