---
title: 'Problem Set #5'
date: "7/31/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE, echo=F}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)

```

**Create an .Rmd file called ps_5.Rmd. Please do not load any packages besides `tidyverse` and (if you want to) `ggthemes`. Please don't forget the instructions and Style Guide:** 

https://rpubs.com/tylersimko/gov50-assignments

## Question 1 

Create an urn in a tibble called `urn` containing 1000 red, blue, and green coins each. `urn` must have two columns: `id` (containing the numbers from 1 to 3000) and `color` (the color of each coin). 

Then, create a new tibble from scratch called `coin_samples` with two columns: `count` which contains only the numbers from 1 to 100, and a list-column called `sampled_coins` which maps `count` and uses the `sample_n` function (not `rep_sample_n`) to sample 10 coins from `urn` for each row.

```{r Q1}

# Simulate an urn with 1000 coins of each color.

urn <- tibble(id = 1:3000, 
              color = c(rep("red", 1000),
                        rep("green", 1000),
                        rep("blue", 1000)))

n <- 10
reps <- 100

# Draw 100 samples of size 10. 
# Note that replacement *between samples* is implicit.
  
coin_samples <- tibble(count = 1:reps,
                       sampled_coins = map(count, ~ sample_n(urn, size = n)))

```

**Final Answer**: `coin_samples` will be a tibble with `r nrow(coin_samples)` rows and `r ncol(coin_samples)` columns: `r colnames(coin_samples)`. 


## Question 2

Even if we didn't know the true answer of all the coins, we could use our random samples to estimate the proportions of each coin color in the urn. `summarize` your `coin_samples` into a tibble with three columns: `prop_red`, `prop_green`, and `prop_blue` with single numeric values in them based on the average proportion of each coin in each sample.

You can do this anyway you like as long as the final answer called `q2` has the correct dimensions, but we recommend:

1. Creating new columns in `coin_samples` called `num_red`, `num_blue`, and `num_green` that use `map` family functions to count the number of red, blue, and green coins within each sample. 
2. Then, you can add three new columns `prop_red`, `prop_blue`, and `prop_green` to calculate the proportion of coins in each sample of that color.
3. Then, you can `summarize` the modified `coin_samples` dataset and overwrite the `prop_red`, `prop_blue`, and `prop_green` columns to take the average of those columns you created in the first step.
4. Save the output into an object called `q2`.

```{r Q2}
q2 <- coin_samples %>% 
  mutate(num_red = map_int(sampled_coins, ~ sum(.$color == "red"))) %>% 
  mutate(num_blue = map_int(sampled_coins, ~ sum(.$color == "blue"))) %>%
  mutate(num_green = map_int(sampled_coins, ~ sum(.$color == "green"))) %>%
  mutate(prop_red = num_red / 10,
            prop_blue = num_red / 10,
            prop_green = num_green / 10) %>%
  summarise(prop_red = mean(prop_red),
            prop_green = mean(prop_green),
            prop_blue = mean(prop_blue))

```

**Final Answer**: `q2` will be a tibble of `r nrow(q2)` rows and `r ncol(q2)` columns made from `coin_samples`. Your final answer must be a tibble with three columns: `prop_red`, `prop_blue`, and `prop_green` with single numeric values in them.

## Question 3

You will write two complementary functions for this problem.

1. Write a function called `scoop` which creates a similar sample as in Question 1 and takes three arguments: `x` for the data, `sample_size` for the size, and `reps` for the number of samples to draw. `scoop` must create and return a new tibble with three columns: `count`, which simply counts from 1 to `reps`, `sampled_coins`, the results from calling `map` on `sample_n` to sample `sample_size` rows from your `x` dataset, and `sample_size`, which is simply equal to the `sample_size` argument you passed. 

You can test this function by running `scoop(x = urn, sample_size = 10, reps = 100)` in your console after creating it. This should make a tibble that looks a lot like `coin_samples` from Question 1. To write `scoop`, try copy-pasting your code from Question 1 and modifying it to create this function.

```{r Q3}

# This function is created for two reasons:
  # 1. It is good to practice writing your own functions
  # 2. To practice using pre-wiitten functions with map.

scoop <- function(x, sample_size, reps) {
  
  # This is essentially identical to coin_samples from 
  # Question 1 with edited object names.
  
  tibble(count = 1:reps, 
         sampled_coins = map(count, ~ sample_n(x, size = sample_size)),
         sample_size = sample_size)
}

```

2. Write a function called `props()` that takes two arguments, `x` and `sample_size`, and calculates the proportion of each color within each sample. This function assumes that `x` is in the format of the output from `scoop`, so you can assume that the list-column wthin `x` is called `sampled_coins`.

This function essentially does steps 1 and 2 from Q2, so you can paste and edit your code from Q2 to create this.

```{r Q3.2}

props <- function(x, sample_size) {
  
  # Group by sample repetition & tally up the widgets by color.
  
  x %>% 
  mutate(num_red = map_int(sampled_coins, ~ sum(.$color == "red"))) %>% 
  mutate(num_blue = map_int(sampled_coins, ~ sum(.$color == "blue"))) %>%
  mutate(num_green = map_int(sampled_coins, ~ sum(.$color == "green"))) %>%
  mutate(prop_red = num_red / sample_size,
            prop_blue = num_blue / sample_size,
            prop_green = num_green / sample_size)
}

```

**Final Answer**: you must define two functions, `scoop` and `props`. Please do not evaluate these functions and print the results to your PDF, although we do encourage you to test them in the console. `scoop` must take three arguments - `x`, `sample_size`, and `reps` that behave as defined above - and return a tibble with three columns - `count`, `sampled_coins`, and `sample_size` exactly as defined above. `props` must take two arguments, `x` and `sample_size`, and return a modified version of `x` with six additional columns (for 9 in total): `num_red`, `num_blue`, `num_green`, `prop_red`, `prop_blue`, and `prop_green`. 

## Question 4

Create a new tibble called `sample_sizes` that has a single column: a vector called `size` which uses the `seq` function to create a vector of all numbers between 10 and 1000 separated by 10 (your resulting vector will have length 100).

Then, using the functions created above, construct a tibble called `mapped_sample` with two columns named `scoop_results` and `props_results` respectively. You will create this by piping `sample_sizes` into a `mutate` call that creates two new columns. The first new column, `scoop_results`, comes from using a `map` function to call `scoop` on the `urn` object with each of your sample sizes 50 times. The second column, `props_results`, comes from calling your new `props` function on each `scoop_results` you made in the previous column. You will need to use the `sample_size` column that `scoop_results` contains.

We recommend creating `sample_sizes` on a smaller vector first (for example, a vector of 10, 20, and 30 alone), create the `mapped_sample` object, and change the `sample_sizes` vector to the full length. 

```{r Q4}

sample_sizes <- tibble(size = seq(10, 1000, by = 10))

mapped_sample <- sample_sizes %>% 
  
  # map(...~ scoop()) iterates over the vector of sample sizes & 
  # takes 50 samples each of that size. The '.' is shorthand for 
  # "each element of the vector we're iterating over" (i.e. `size`).
  # map(...~ props()) does the same with calculating proportions.
  # However, we need to specify the .$sample_size argument here.
  
  mutate(scoop_results = map(size, 
                             ~ scoop(x = urn, sample_size = ., reps = 50))) %>% 
  mutate(props_results = map(scoop_results,
                             ~ props(x = ., sample_size = .$sample_size)))

```

**Final Answer**: `mapped_sample` must be a new tibble, created from `sample_sizes`, with `r nrow(mapped_sample)` rows and `r ncol(mapped_sample)`: `r colnames(mapped_sample)`. Columns `scoop_results` and `props_results` must be list-columns.

## Question 5

Use `mapped_sample` to calculate the standard deviation of each color in all of the draws for each sample size. You will need to start with `mapped_sample` and pipe it into the creation of three new columns which estimate the standard deviation (try the `sd()` function) of each "prop" column created in the previous question. Call these new columns `props_red_sd`, `props_green_sd`, and `props_blue_sd`. Save this into an object called `sd_results`. 

```{r Q5}

# map_dbl() takes the output of prop() for each sample size & allows
# pull() to extract the named column, which is then piped into the calculation of SD.
# SD is the spread in the proportions of a certain color across samples of each size.
# Note that map_dbl() always returns a double (a real number).

# The .$ syntax would work too:
# mapped_sample %>% mutate(props_red_sd = map_dbl(props_results, ~ sd(.$prop_red)))

sd_results <- mapped_sample %>% 
  mutate(props_red_sd = map_dbl(props_results, ~ pull(., prop_red) %>% sd())) %>% 
  mutate(props_green_sd = map_dbl(props_results, ~ pull(., prop_green) %>% sd())) %>% 
  mutate(props_blue_sd = map_dbl(props_results, ~ pull(., prop_blue) %>% sd()))

```


**Final Answer**: `sd_results` will be a tibble of `r nrow(sd_results)` rows and `r ncol(sd_results)` columns: `r colnames(sd_results)`, where the new columns `props_red_sd`, `props_green_sd`, and `props_blue_sd` each contain numeric values.

## Question 6

Design a plot that visualizes the relationship between sample size and standard deviation for each of the three colors by using `sd_results`. Your plot must:

1. Clearly distinguish between the standard deviation estimates for the red, blue, and green colors. Below, we do that by plotting each line separately in a different color. Your plot does not have to use lines, but it can. 
2. Use better legend names than "props_blue_sd", "props_red_sd", and "props_green_sd". Hint: use a scale function!
3. Use all of the `size` and `std_dev` values from `sd_results`. That is, for example, do not filter out to indvidual size values.
4. Use a theme other than the default `ggplot` theme.

Here is an example, created by pivoting `sd_results` into a 300 x 3 datset. Your plot **does not** need to look anything like this, though it does meet all of the criteria defined above.

```{r Q6}

sd_results %>%
    select(-scoop_results, -props_results) %>%
    pivot_longer(cols = props_red_sd:props_blue_sd, 
                 names_to = "proportion",
                 values_to = "std_dev") %>%
  ggplot(aes(x = size, y = std_dev, color = proportion)) + 
  geom_line() + 
  theme_bw() + 
  scale_color_manual(values = c("#0A2472", "#D8315B", "#26C485"),
                     breaks = c("props_blue_sd", "props_red_sd", "props_green_sd"),
                     labels = c("Blue", "Red", "Green"),
                     name = "Coin Color") + 
  labs(title = "Relationship Between Sample Size and Standard Deviation",
       x = "Sample Size",
       y = "Estimated Standard Deviation")

```
