---
title: "Week 9, Day 1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(PPBDS.data)
library(rstanarm)
library(tidyverse)

# We will use the `kenya` dataset from PPBDS.data. This dataset is from Harris,
# Kamindo and Van der Windt (2020): "Electoral Administration in Fledgling
# Democracies:Experimental Evidence from Kenya." The authors worked with
# Kenya's electoral commission in 1,674 communities by assigning polling
# stations to either a control group or to one of five other treatments.

# For simplicity we will only consider polling stations assigned to either
# "control" or to "local". The latter means a treatment in which election
# officials visited the community's polling station for two days, thereby making
# it easier for people to register.

# The outcome variable, `reg_chg`, is the change in voter registration in each
# community. `poverty` is the percentage of the community below the poverty
# level. `distance` is the number of kilometers between the polling station and
# the central political office (where registrations normally take place).
# `pop_density` is a measure of population density around the polling station.


week_9 <- kenya %>% 
  rename(reg_chg = reg_byrv13) %>% 
  filter(treatment %in% c("control", "local")) %>% 
  droplevels() %>% 
  select(reg_chg, treatment, poverty, 
         distance, pop_density)
```


## Scene 1

**Prompt:** In addition to the outcome variable and the treatment, we have 3 statistics for the communities in the sample --- poverty, distance and population density. See code comments above for details. Never hurts to do a little exploratory data analysis. Look at the data!

* Make a scatter plot of `poverty` on the y-axis and `distance` in the x-axis. Include a `geom_smooth()` with `method` set to "lm" and `se` equal to FALSE. Write a sentence to answer these questions:

  + Does poverty cause distance? Or does distance cause poverty? Or both? Or neither?
  + How accurate is the model beyond 90 kilometers?  
  + Eye-balling the chart, how different is the poverty rate between two polling stations whose distance differs by 50 kilometers?
  
* Fit a `stan_glm()` model which relates poverty to distance. Print a summary. (Hint: In order to see the coefficient of `distance`, you will probably need to use the `digits` argument to `print()`.) Interpret the meaning of the two important parameters.  


```{r sc1-a}
week_9 %>% 
  ggplot(aes(x = distance, y = poverty)) +
    geom_point() +
    geom_smooth(method = "lm", se = FALSE) +
    labs(title = "Poverty and Distance at Kenyan Polling Stations",
         subtitle = "Poorer communities are further away from registration offices",
         x = "Distance (km)",
         y = "Poverty Measure",
         caption = "Data from Harris, Kamindo and Van der Wind (2020)") + 
  ylim(c(0,1))
```


```{r sc1-b}
fit_1 <- stan_glm(poverty ~ distance,
                  data = week_9,
                  refresh = 0)

print(fit_1, digits = 4)
```

## Scene 2

**Prompt:** Let's make a model with `reg_chg` as the outcome variable and with treatment, poverty, distance and pop_density as the predictors. (Using outcome/predictors is the typical nomenclature in statistics. In Economics, we would refer to dependent/independent variables.)

* Provide the math for this model.

* Estimate the model. Assign the fitted model to `fit_2`. Interpret the important parameters of the model.

$$ reg\_chg_i = \beta_0 + \beta_1 local + \beta_2 poverty_i + \\
                \beta_3 distance_i + \beta_4 pop\_density_i +  + \epsilon_i
$$


```{r sc2-a}
fit_2 <- stan_glm(data = week_9,
                  formula = reg_chg ~ treatment + poverty + distance + pop_density,
                  refresh = 0)

print(fit_2, detail = FALSE, digits = 4)
```