---
title: "Week 6, Day 1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)

# Build an urn with 500 red beads, 490 blue beads, and 10 yellow beads. The urn
# should have two variables: ID and color. Mix up the beads in the urn randomly.
# That is, we don't want all the red beads to have ID 1 through 500 and so on.
# Sort the urn by ID.

beads <- c(rep("red", 500), 
           rep("blue", 490), 
           rep("yellow", 10))

urn <- tibble(color = beads) %>% 
  sample_frac() %>% 
  mutate(ID = 1:1000) %>% 
  select(ID, color)

# Using the book, where a very similar example is given, is highly recommended.
# But, this example is trickier than the book because you have to mix up the
# beads before you assign the `ID` values. `sample_frac()` is a handy function.
# Check out other functions in the sample_* family. I keep thinking that there
# is a better function to use, something which only mixes stuff up, but I could
# not find it. Perhaps a simpler approach:

urn <- tibble(ID = 1:1000,
              color = sample(beads))

# If you don't tell sample() how many you want, it just gives you back
# everything, reordered.
              
```

We are learning about sampling this week. We are taking the theory of Chapter 5 and applying it in a not-too-toyish example. There is a single, true, unknown parameter. What is your posterior distribution for that parameter? Once you have that posterior, how can you use it to make forecasts about the future? 


## Scene 1

**Prompt:** Without using rep_sample_n(), sample 30 beads at random from the urn five times. The resulting tibble should have two columns: `replicate` (with a value of 1 to 5) and `samp` (a list column in which each item is a tibble with 30 rows sampled, with replacement, from `urn`). Each tibble in `samp` will have the same two columns as `urn`.

```{r sc1}

# First, make an empty tibble with 5 "slots" for samples
tibble(replicate = 1:5) %>% 
  
  # Make a list-column and take a sample of size 30 from urn with each of those.
  # replace = TRUE means we will put back each bead as we sample it, instead of 
  # storing the ones we've already sampled aside.
  
  mutate(samp = map(replicate, ~ sample_n(urn, 30, replace = TRUE))) 

```


## Scene 2

**Prompt:** What is the probability that, if you sample 50 beads at random, there will be more red than blue beads in the sample?


```{r sc2}
tibble(replicate = 1:1000) %>% 
  
  # Very similar to the code above! Just changing some constants.
  
  mutate(samp = map(replicate, ~ sample_n(urn, 50, replace = TRUE))) %>% 
  
  # We now use the list-column we created, samp, to make integer values.
  # That's why we use map_int and not map.
  
  mutate(numb_red = map_int(samp, ~ sum(.$color == "red"))) %>% 
  mutate(numb_blue = map_int(samp, ~ sum(.$color == "blue"))) %>% 
  
  # Now we can simply compare two integer values
  
  mutate(red_wins = ifelse(numb_red > numb_blue, TRUE, FALSE)) %>% 
  summarize(prob = sum(red_wins)/n())


```

## Scene 3

**Prompt:** When performing inference, it is often useful to ask: What are the range of things which might happen if X is true? We saw that in several examples in chapter 5. Assume that a person is sick, how likely is she to test positive for the disease? Assume that there is one white marble in the bag, how likely are we to draw zero white marbles out of three?

Let's follow that same intuition in the case of the urn. We have an urn with 1,000 beads. Assume that there are 400 red beads. (This is not true! In fact, we know that there are 500 red beads in the urn. But make the assumption anyway.)

* Using `rbinom()` calculate the number of red beads we might draw from the urn with a paddle of size 25. Hint: `n` is 1 in this case. What does it mean if we set `n` equal to 10?

* Create a tibble with two columns: `numb_red_urn` and `numb_red_paddle`. Initially, the tibble has one row, with `numb_red_urn` set to 400, as in the previous question, and `numb_red_paddle` being the number of red drawn using `rbinom`.

* Extend that tibble to have 7 rows, with `numb_red_urn` set to 100, 200, ..., 700. Make sure that the results in `numb_red_paddle` make sense. `seq()` and `map_int()` are useful.

* Still using a tibble with those 7 rows, call `rbinom()` with `n = 10`. (This may require other changes in your code.) Turn the resulting tibble, in which `numb_red_paddle` is a list column into an easier-to-work-with tibble in which `numb_red_paddle` is an integer. Hint: `unnest()`. Interpret the resulting object.

```{r}

# Using `rbinom()` calculate the number of red beads we might draw from the urn with a paddle of size 25. Hint: `n` is 1 in this case. What does it mean if we set `n` equal to 10?

rbinom(n = 1, size = 25, prob = 400/1000)

# This works for one row.

tibble(numb_red_urn = 400) %>% 
  mutate(numb_red_paddle = rbinom(n = 1, size = 25, prob = numb_red_urn/1000))

# But not for multiple rows! 
tibble(numb_red_urn = seq(100, 700, by = 100)) %>% 
  mutate(numb_red_paddle = rbinom(n = 1, size = 25, prob = numb_red_urn/1000))

# For that, you need map.
tibble(numb_red_urn = seq(100, 700, by = 100)) %>% 
  mutate(numb_red_paddle = map_int(numb_red_urn, ~ rbinom(n = 1, size = 25, prob = ./1000)))

# More experiments and unnesting for convenience. 
tibble(numb_red_urn = seq(100, 700, by = 100)) %>% 
  mutate(numb_red_paddle = map(numb_red_urn, ~ rbinom(n = 10, size = 25, prob = ./1000))) %>% 
  unnest(numb_red_paddle)
```

## Original scenes 3-5 coming on Thursday!
